---
title: "RStudio Project Report"
author: 'Author: The Vinh (Vin) Ha'
output: pdf_document
---

### Study on Traffic dataset:

**a) Correlation plot and matrix, comments on relationships of predictors and response:**

```{r}
traffic = read.csv('data/traffic.csv', header = TRUE)
pairs(traffic, panel = panel.smooth)
```

```{r}
cor(traffic)
```

- The response variable ***spi*** has a strong negative relationship with the predictor ***transport***; a weak negative relationship with the predictor ***road***; a strong positive relationship with the predictor ***weather***; and no obvious relationship with both the predictors ***fuel*** and ***wind***.

- There does not seem to be a relationship between the predictors themselves.

**b) Fit full model and estimate the impact of *weather* on *spi* with 95% CI:**

```{r}
M1 = lm(spi ~ ., data = traffic)
summary(M1)
```

```{r}
summary.M1 = summary(M1)
sqrt(diag(summary.M1$cov.unscaled * summary.M1$sigma^2))[4]
qt(1 - 0.05/2,56)
```

**The require CI is:**

$$
\hat{\beta}_{\text{weather}} \pm t_{n-p,1-\frac{\alpha}{2}}\text{se}(\hat{\beta}_{\text{weather}})\\ 
= \hat{\beta}_{\text{weather}} \pm t_{56,0.975}\text{se}(\hat{\beta}_{\text{weather}}) \\ 
= 4.2456 \pm 2.003241 \times 0.4472731 \\
= (3.349604, 5.141596)
$$

That is, we are 95% confident that for every percentage increase in relative ***weather***, the ***spi*** concentration will increase between **3.349604** and **5.141596** on average.

**c) Conduct F-test for overall regression and examine relationship between predictors and response:**

**Theorical Model is:**

$$Y_i = \beta_0 + \beta_1 X_{i1} + \beta_2 X_{i2} + \beta_3 X_{i3} + \beta_4 X_{i4} + \beta_5 X{i5} + \epsilon_i, \quad i = 1, 2, \ldots, n
$$

- $Y_i$ is the response variable ***spi***
- $X_{ij}$ is th are the predictors variables for the i-th observation:
  * $X_{i1}$ = annual mean ***transport*** of test locations
  * $X_{i2}$ = annual mean ***road*** of test locations
  * $X_{i3}$ = annual mean ***weather*** of test locations
  * $X_{i4}$ = annual mean ***fuel*** of test locations
  * $X_{i5}$ = annual mean ***wind*** of test locations
- $\epsilon \sim \mathcal{N}(0, \sigma^2)$ denotes the random variation with constant variance

**Now we conduct the F-test:**
  
  - Hypotheses: $H_0: \beta_1 = \ldots = \beta_5  = 0 \text{ vs }  H_1: \text{not all } \beta_i = 0, \text{ for } i = 1, 2, \ldots, 5$
  
  - Standard R output ANOVA table:
  
```{r}
anova(M1)
```

- Note the $Regression SS = 4742.6 + 1992.7 + 8651.9 + 258.1 + 58.2 = 15703.5$
- Therefore the Mean Squared Reg = Reg SS / Reg df = $15703.5 / 5 = 3140.7$
- Test statistics: $F_{obs} = MS_{Reg} / MS_{Res} = 3140.7 / 98.26094 = 31.96285$
- The null distribution for the test statistics is: $F_{5,56}$
- P-value: $P(F_{5,56} \geq 31.96285) = 0 = 3.0386681751212669e-15 < 0.05$
- As the P-value is small:
  - (Statistical) There is enough evidence to reject $H_0$
  - (Contextual) There is significant linear relationship between ***spi*** and at least one of the 4 predictor variables.
  
**d) For the diagnostics:**

```{r}
par(mfrow = c(1,2))
plot(M1, which = 1:2)
par(mfrow = c(1,3))
plot(resid(M1) ~ transport, data = traffic, xlab = "transport", ylab = "Residuals")
abline(h = 0, lty = 2)
plot(resid(M1) ~ road, data = traffic, xlab = "road", ylab = "Residuals")
abline(h = 0, lty = 2)
plot(resid(M1) ~ weather, data = traffic, xlab = "weather", ylab = "Residuals")
abline(h = 0, lty = 2)
par(mfrow = c(1,2))
plot(resid(M1) ~ fuel, data = traffic, xlab = "fuel", ylab = "Residuals")
abline(h = 0, lty = 2)
plot(resid(M1) ~ wind, data = traffic, xlab = "wind", ylab = "Residuals")
abline(h = 0, lty = 2)
```

- The quantile plot of residuals look approximately linear, so the normality assumption for residuals is appropriate

- There is no obvious pattern in any of the residual plots so it appears the linearity and constant variance assumptions of the multiple linear model are 

**e) Find R2:**

- Here $R^2$ = 0.741 = 74.1%, which is a goodness of fit metric. It means 74,1% of the variation in ***spi*** is explained by the full linear regression model.

**f) Find best regression model:**

```{r}
summary(M1)
```

- ***wind*** has the highest P-value so remove it first

```{r}
M2 = update(M1, . ~ . - wind)
summary(M2)
```

- ***fuel*** still has large P-value so remove it

```{r}
M3 = update(M2, . ~ . - fuel)
summary(M3)
```

- In model M3 although all predictors are significant, the value of $R^2$ adjust decrease, which indicates that the recently removed predictor ***fuel*** has contribution to the model. Therefore, we will stop at the final model M2.  

$$\hat{Y} = 61.16096 - 2.12565X_1 - 2.43721X_2 + 4.25645X_3 - 3.68533X_4$$

**g) Explain R2 and R2 adjust:**

- The R2 goodness of fit metric always decreases/increases when a predictor is removed/added from/into the model. The adjusted R2 has a penalty for the number of predictors in the model. So it will sometimes increase when a predictor is removed. In this case, from the full to final model, the R2 decreases from 74.1% to 73.8% but the adjusted R2 increases from 71.7% to 71.9%.

### Study on Cake dataset:

**a) Balanced study checking and explain:**

- A study is balanced if there are equal number of replicates across all the levels factors in the study.

- We can check replicates by: 

```{r}
cake  = read.csv('data/cake.csv', header = TRUE, stringsAsFactors = TRUE)
table(cake[, c("Recipe", "Temp")])
```
- From the above we can see that the design is balanced with an equal number of replicates for each combination of levels of the two factors.

**b) Preliminary graphs that investigate different features of the data:**

```{r}
boxplot(Angle ~ Recipe + Temp, data = cake)
with(cake, interaction.plot(Temp, Recipe, Angle))
```

- From the boxplot, we can see that the assumption of equal variance among levels seems approximately valid due to the similar box sizes.

- From the interaction plot we can see there is non-parallel lines  for the means of each group at different levels of the independent variables, this indicates a significant interaction effect between the two independent variables.

**c) Interaction model:**

- The Two-Way ANOVA model with interaction is:

$$Y_{ijk} = \mu + \alpha_i + \beta_j + \gamma_{ij} + \epsilon_{ijk}$$

- The parameters are:
  - $Y_{ijk}$: the reponse breaking angle of the cake
  - $\alpha_i$: the ***Recipe*** effect, there are 3 levels: A, B, C
  - $\beta_j$: the ***Temp*** effect, there are 6 levels: 175C, 185C, 195C, 205C, 215C, 225C
  - $\gamma_{ij}$: the interaction effect between ***Recipe*** and ***Temp***
  - $\epsilon_{ijk} \sim \mathcal{N}(0, \sigma^2)$: the unexplained variation

**d) Study the effect of *Recipe* and *Temp* on *Angle*:**

- We will conduct a hypotheses test:

$$
H_0: \gamma_{ij} = 0 \text{ for all } i, j
\\H_1: \text{at least one } \gamma_{ij} \neq 0
$$

- Now we fit the interaction model:

```{r}
cake.int = lm(Angle ~ Recipe * Temp, data = cake)
anova(cake.int)
```

- Since the P-value is 0.775632 > 0.05. we do not have enough evidence to reject $H_0$. We can see that the interaction term is insignificant. Therefore, this is not the final model yet.

- We should validate the interaction model with diagnostic plots:

```{r}
par(mfrow = c(1:2))
plot(cake.int, which = 1:2)
```

  - The residuals are close to linear in the QQ-plot, and so the normal assumption should be valid. The residual plot seems to show equal spread around the fitted values and so the constant variance assumption is also appropriate.
  
**e) Repeat test analysis for the main effects:**

- We will conduct hypotheses tests for main effects:
$$
H_0: \alpha_i = 0 \text{ for all } i \ \ \ \ \ \ \text{against} \ \ \ \ \ \  H_1: \alpha_i \neq 0 \\
H_0: \beta_j = 0 \text{ for all } j \ \ \ \ \ \ \text{against} \ \ \ \ \ \  H_1: \beta_j \neq 0
$$

- Now we refit the model without interaction term:

```{r}
cake.main = lm(Angle ~ Recipe + Temp, data = cake)
anova(cake.main)
par(mfrow = c(1,2))
plot(cake.main, which = 1:2)
```

- The result of ANOVA table shows that the P-values for both ***Recipe*** and ***Temp*** are smaller than 0.05, which help reject all $H_0$ above and indicates that the main effects are significant.

- Again, the residuals are close to linear in the QQ-plot, and so the normal assumption should be valid, and the residual plot seems to show equal spread around the fitted values and so the constant variance assumption is also appropriate.

**f) Conclusion on effect:**

- Overall, the effect of ***Recipe*** and ***Temp*** on the quality of cakes ***Angle*** are not depend on each other since the interaction term is insignificant. However, these effects separately have significant impact on the response ***Angle***, that is, ***Angle*** seems to be higher with ***Recipe*** A and B, within the ***Temp*** ranges of 195C, 215C, 225C.

- We can interpret the main effects separately since there is no significant interaction effect them.




